Bayesian Detection of Bias in Peremptory Challenges Using Historical Strike Data

### 2023-01-05 analysis/paper_draft3.Rmd
minor edits

```
In this paper, we extend a Bayesian approach to _Batson_ and similar challenges [@Kadane2021; @Kadane2018a; @Kadane2018b] to incorporate historical strike data in a manner that allows us to explicitly incorporate the degree of incompatibility between current and historical trials. In this approach, we use two steps. In step one, we specify a model of the peremptory-strike process in the court of interest that includes a bias parameter to which we assign an initial prior distribution. We use an attorney's strike data from past trials, as generated by that same peremptory-strike process, to estimate a posterior distribution for that bias parameter. We use the power prior [@Chen2000; @Ibrahim2015] to control how much the strike history from past cases affects the posterior distribution of the bias parameter. In step two, we estimate the posterior distribution of the bias parameter from the current case, using the posteri-or distribution developed with the historical information and the power prior.

The use of the power prior in this context allows lawyers to explicitly control over the assumed degree of incompatibility between the current and historical trial data. Specifically, lower values (closer to 0) of the power prior parameter, $\alpha$, should be selected when incompatibility is expected to be high, and higher values (closer to 1) of the power prior parameter selected when incompatibility is low. Accordingly, the lawyer bringing the _Batson_ challenge chooses the value of $\alpha$ and must be able to explain and defend that choice as appropriately cooresponding to the likely degree of incompatibility between the current strike data and historical strike data available in the particular case.

>>>>

```
In this paper, we extend a Bayesian approach to _Batson_ and similar challenges [@Kadane2021; @Kadane2018a; @Kadane2018b] to incorporate historical strike data in a manner that allows us to explicitly incorporate the degree of incompatibility between current and historical trials. In this approach, we use two steps. In step one, we specify a model of the peremptory-strike process in the court of interest that includes a bias parameter to which we assign an initial prior distribution. We use an attorney's strike data from past trials, as generated by that same peremptory-strike process, to estimate a posterior distribution for that bias parameter. We use the power prior [@Chen2000; @Ibrahim2015] to control how much the strike history from past cases affects the posterior distribution of the bias parameter. In step two, we estimate the posterior distribution of the bias parameter from the current case, using the posterior distribution developed with the historical information and the power prior.

The use of the power prior in this context allows lawyers to explicitly control the assumed degree of incompatibility between the current and historical trial data. Specifically, lower values (closer to 0) of the power prior parameter, $\alpha$, should be selected if incompatibility is expected to be high. Higher values (closer to 1) of the power prior parameter should be selected if incompatibility is expected to be low. Accordingly, the lawyer bringing the _Batson_ challenge, after choosing the value of $\alpha$, must be able to defend that choice as appropriately cooresponding to the likely degree of incompatibility between the current and historical strike data in the particular case.
